---
date: "05/27/2026" 
author: "Kierra Miller"
title: "Homework 3_ENVS 193DS"
format: html
---

# Part 1. Set Up 

```{r}
# Homework Set Up and install packages
library(tidyverse)
library(here)
library(janitor)
library(readxl)


kelp <- read_csv(here("data", "temp-kelp.csv"))

```

# Part 2. Problems
## Problem 1. Giant Kelp Fronds

### a. Appropriate test 

The appropriate tests would include Pearson correlation and linear regression. Pearson correlation measures the strength and direction of the relationship between temperature and kelp elongation rate, while linear regression models how kelp growth changes with temperature

### b. Visualization 

```{r}
ggplot(kelp, aes(x = temp_c, y = kelp_elong)) +
  geom_point(color = "darkgreen") +
  geom_smooth(method = "lm", color = "blue") +
  labs(
    x = "Ocean temperature (°C)",
    y = "Kelp elongation rate (cm/day)",
    title = "Kelp growth vs ocean temperature"
  ) +
  theme_minimal()

```


### c. Check your assumptions and run test 

```{r}
# check relationship 
plot(kelp$temp_c, kelp$kelp_elong)

# check distribution 
hist(kelp$kelp_elong)

```
I checked linearity using a scatterplot and assessed the distribution using a histogram. The relationship appears approximately linear with no significant outliers, and the data appear reasonably normally distributed. 

```{r}
# Run regression 
model <- lm(kelp_elong ~ temp_c, data = kelp)
summary(model)
```



### d. Results communication 



### e. Test implications 




### f. Double check work









```

