---
date: "05/27/2026" 
author: "Kierra Miller"
title: "Homework 3_ENVS 193DS"
format: html
---

# Part 1. Set Up 

```{r}
# Homework Set Up and install packages
library(tidyverse)
library(here)
library(janitor)
library(readxl)

# creating object and reading in personal data set 
my_data <- read_csv("personal-data (2).csv",
              show_col_types = FALSE,
                    col_names = TRUE) 
# clean names
my_data_clean <- my_data %>%
  clean_names()

kelp <- read_csv(here("data", "temp-kelp.csv"))

```

# Part 2. Problems
# Problem 1. Giant Kelp Fronds

## a. Appropriate test 

The appropriate tests would include Pearson correlation and linear regression. Pearson correlation measures the strength and direction of the relationship between temperature and kelp elongation rate, while linear regression models how kelp growth changes with temperature

## b. Visualization 

```{r}
ggplot(kelp, aes(x = temp_c, y = kelp_elong)) +
  geom_point(color = "darkgreen") +
  geom_smooth(method = "lm", color = "blue") +
  labs(
    x = "Ocean temperature (°C)",
    y = "Kelp elongation rate (cm/day)",
    title = "Kelp growth vs ocean temperature"
  ) +
  theme_minimal()

```


## c. Check your assumptions and run test 

```{r}
# check relationship 
plot(kelp$temp_c, kelp$kelp_elong)

# check distribution 
hist(kelp$kelp_elong)

```
I checked linearity using a scatterplot and assessed the distribution using a histogram. The relationship appears approximately linear with no significant outliers, and the data appear reasonably normally distributed. 

```{r}
# Run regression 
model <- lm(kelp_elong ~ temp_c, data = kelp)
summary(model)
```

## d. Results communication 

A linear regression was used to test the relationship between temperature and kelp growth. The results show a significant negative relationship (p = 1.366e-05, $R^2$ = 0.473), indicating that as temperature increases, kelp elongation rate decreases. Indicating that temperature does significantly predict kelp elongation rate. 

## e. Test implications 

These results suggest that increasing ocean temperatures may impact kelp growth rates. Since the relationship appears to be negative, warming oceans could reduce kelp productivity, affecting ecosystem health and function. This information can help guide conservation and climate adaptation and management strategies. 


## f. Double check work

```{r}
# Pearson correlation 
cor.test(kelp$temp_c,kelp$kelp_elong)

```

Both Pearson correlation and linear regression led to the same reasonable conclusion about the relationship between temperature and kelp elongation rate. While correlation measures strength and direction, regression additionally quantifies the predictive relationship between variables. 

# Problem 2. Personal Data 

## a. Update plots 

```{r}
# Updated visualization plots from Homework 2 

# updating Plot 1 (Categorical) 
ggplot(my_data_clean, aes(x = class_location, y = ounces_consumed, fill = class_location)) +
  
  # jitter points
  geom_jitter(width = 0.15, size = 3, alpha = 0.8) +
  
  # custom colors
  scale_fill_manual(values = c("orchid","skyblue","gold","green","purple","tomato")) +
  
  # labels
  labs(
    title = "Caffeine Consumption by Class Location",
    subtitle = "Latest observation: May 28, 2026",
    x = "Class Location",
    y = "Caffeine Consumed (oz)"
  ) +
  
  # clean theme
  theme_classic() +
  
  # formatting
  theme(
    legend.position = "none",
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

```{r}
# updating Plot 2 (Continuous)

ggplot(my_data_clean, aes(x = number_of_classes, y = ounces_consumed)) +
  
  # scatter points
  geom_point(size = 3, color = "darkorange") +
  
  # x-axis formatting
  scale_x_continuous(breaks = c(1, 2, 3)) +
  
  # labels
  labs(
    title = "Caffeine Consumption and Class Load",
    subtitle = "Latest observation: May 28, 2026",
    x = "Number of Classes",
    y = "Caffeine Consumed (oz)"
  ) +
  
  # clean theme
  theme_minimal() +
  
  # reduce clutter
  theme(
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    plot.title = element_text(size = 14, face = "bold")
  )

```

## b. captions 

Figure 1. Caffeine consumption (oz) varies across class locations. Each point represents an observation day, with jitter applied to reduce overlap. Consumption appears to vary by location, suggesting that where classes are held may influence caffeine intake patterns. 

Figure 2. Caffeine consumption (oz) in relation to number of classes attended per day. Each point represents a single day. While there is some variability, higher class loads may be associated with increased caffeine consumption. 













```









